An Implementation of the Active Contours without Edges model and the Logic Framework for Active contours on multi-channel images

### 2010-12-13 trunk/cs870/project/Report/Paper.tex
trying to fix missing references

\title{An Implementation of the Active Contours without Edged model and the Logic Framework for Active contours on multi-channel images}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg1.png}
\caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
Bottom: evolving segmentation over time until the objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, $\mu =
0.01$, no reinitialization, cpu = 2.9s, 7 iterations.}
\label{fig:cv_eg1}
\end{figure*}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg7.png}
\caption{Inability to detect low intensities that have not much variation from their background.  Top: the evolving curve (in red) over time where the first
image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
$\mu =0.01$, no reinitialization, cpu = 1.35 s, took 6 iterations.}
\label{fig:cv_eg7}
\end{figure*}

@@

@@

There are mainly three parameters in this model that can be varied: $\mu, \lambda_{1}, \lambda_{2}$. We would usually want both $\lambda_{1}$ and $\lambda_{2}$
to be equal to 1 to indicate that we care both about the difference of intensities inside and outside. A smaller weighting for one of them would mean that we
would ignore some of the variances in intensities in this area. A larger weighting means that we are magnifying the difference in intensities in that area
(i.e. we want to detect any minor changes). We show an example of these variations in Figure~\ref{fig:cv_eg12} where we vary
$lambda_1$ and $\lambda_2$. When we set $\lambda_1 = 0.2$ and $\lambda_2 = 1$, less details within the interior of the brain is detected. When
we set $\lambda_1 = 2$ and $\lambda_2 = 1$, we can see that more details within the brain are detected. When we set $\lambda_1 = 5$, and $\lambda_2 = 0.01$ (a
very extreme case), we can see that parts of outside boundaries are not well detected, while many details are detected within the brain.

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{cv_eg12.png}
\caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour. The figure shows the segmentation in each case.}
\label{fig:cv_eg12}
\end{figure*}

@@

For the Sandberg-Chan model, the successful detection of boundaries criteria is implicitly

@@

@@

\subsubsection*{Union}

\subsubsection*{Intersection}

@@

>>>>

\title{An Implementation of the Active Contours without Edges model and the Logic Framework for Active contours on multi-channel images}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=11cm]{cv_eg1.png}
\caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
Bottom: evolving segmentation over time until the objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, $\mu =
0.01$, no reinitialization, cpu = 2.9s, 7 iterations.}
\label{fig:cv_eg1}
\end{figure*}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=10cm]{cv_eg7.png}
\caption{Inability to detect low intensities that have not much variation from their background.  Top: the evolving curve (in red) over time where the first
image shows the initial
contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
$\mu =0.01$, no reinitialization, cpu = 1.35 s, took 6 iterations.}
\label{fig:cv_eg7}
\end{figure*}

@@

\begin{figure}[t!]
\centering
\includegraphics[width=6cm]{sc_unionsimple.png}
\caption{Simple union example. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 1.32 s, 
iterations = 2.}
\label{fig:sc_unionsimple}
\end{figure}

\begin{figure}[t!]
\centering
\includegraphics[width=6cm]{sc_intersectionsimple.png}
\caption{Simple intersection example. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 1.83 s, 
iterations = 2.}
\label{fig:sc_intersectionsimple}
\end{figure}

@@

There are mainly three parameters in this model that can be varied: $\mu, \lambda_{1}, \lambda_{2}$. We would usually want both $\lambda_{1}$ and $\lambda_{2}$
to be equal to 1 to indicate that we care both about the difference of intensities inside and outside. A smaller weighting for one of them would mean that we
would ignore some of the variances in intensities in this area. A larger weighting means that we are magnifying the difference in intensities in that area
(i.e. we want to detect any minor changes). We show an example of these variations in Figure~\ref{fig:cv_eg12} where we vary
$\lambda_1$ and $\lambda_2$. When we set $\lambda_1 = 0.2$ and $\lambda_2 = 1$, less details within the interior of the brain is detected. When
we set $\lambda_1 = 2$ and $\lambda_2 = 1$, we can see that more details within the brain are detected. When we set $\lambda_1 = 5$, and $\lambda_2 = 0.01$ (a
very extreme case), we can see that parts of outside boundaries are not well detected, while many details are detected within the brain.

\begin{figure}[t]
\centering
\includegraphics[width=7cm]{threech.png}
\caption{Logic operations performed on three channels. The figure shows the final segmentation in each case. Size = 200 * 200. $\phi_{0}(x,y) = - \sqrt{(x -
100)^2 + (y - 100)^2)} + 50$, $\mu = 0.1$, no reinitialization. For union: cpu = 3.92 sec and iterations = 6. For intersection: cpu = 3.11 sec and iterations =
5. For intersection then union: cpu = 1.24 and iterations = 4.}
\label{fig:threech}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=7cm]{cv_eg12.png}
\caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour. The figure shows the final segmentation in each case.}
\label{fig:cv_eg12}
\end{figure}

@@

For the Sandberg-Chan model, the successful detection of boundaries criteria is implicitly included in the logic operations functionality. We therefore, show
the other criteria. Unless otherwise specified, $\mu = 0.1$ and $\lambda = 255*255$ in the experiments below.

\subsubsection*{Logic Operators: union and intersection}

To ensure that the union operator is working properly, we tried it on a very simple example of two rectangles shown in Figure~\ref{fig:sc_unionsimple}. The
union of the two rectangles was successfully detected. Similarly, Figure~\ref{fig:sc_intersectionsimple} shows the intersection for the same two simple
rectangles.

Figure~\ref{fig:donutlogic} shows a more complicated example for both the intersection and union with the donut image shown in the previous section. We used
two versions of the donuts with each one having a different part occluded. The results of the union and intersection operators are shown in the figure.
Additionally, this example shows that the Sandberg-Chan model can still successfully detect holes despite the addition of the logic operations.

A practical usage of this logic operations framework is to recover missing parts from different images, and combine them to produce a more complete image. We
show this in Figure~\ref{fig:family} which shows two different version of a family image with a child missing in each image. When the union is applied to these
two channels, the complete family image is recovered.

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{donutlogic.png}
\caption{Performing union and intersection on images with holes. Size = 300 * 300. $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
75$, no reinitialization. For union: cpu = 8.63 sec and iterations = 10. For intersection: cpu = 7.68 sec and iterations = 10.}
\label{fig:donutlogic}
\end{figure*}

\begin{figure*}[t]
\centering
\includegraphics[width=12cm]{flowers.png}
\caption{Ability to detect the full contour on a colored image with low variation between foreground and background by taking the union of the RGB channels.
Full contour detected as opposed to Figure~\ref{fig:cv_eg7}. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, no reinitialization,
size = cpu = 1.73 sec, iterations = 7.}
\label{fig:flowers}
\end{figure*}

@@

Similar to the previous model, we needed to ensure that the position of the initial curve does not alter our results. Figure~\ref{fig:sc_position} shows
the segmentation results obtained for the same image used in Figure~\ref{fig:sc_unionsimple}, but with a different position for the initial curve.

\begin{figure}[t]
\centering
\includegraphics[width=7cm]{sc_position.png}
\caption{Same example shown in Figure~\ref{fig:sc_unionsimple}, but with different initial curve position. Initial curve does not overlap with either objects.
Same segmentation is obtained. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 4.43s, 4
iterations.}
\label{fig:sc_position}
\end{figure}

@@

To be able to perform well in noisy images, $\lambda$ must be decreased in order to ignore the noise. We added 'Salt \& pepper' noise with a variation of 0.1 to
the images shown in Figure~\ref{fig:sc_noisy}, and tried to perform the union. In the first case shown, we set $\lambda = 255*255$ which is the normal setting
used for clean images, and in the second case, we set $\lambda = 25$. We show the final segmentation for both cases. In the first case, the noise was detected,
while in the second case, the noise was ignored since $\lambda$ was small. This shows that we also satisfy the response to parameter settings criteria since
$\lambda$ was the only variation mentioned by the authors in the original paper~\cite{sandberg2005logic}. We note that the contour evolved much more slowly
with a smalled $\lambda$, and thus took much more CPU time, and a larger number of iterations.

\begin{figure}[t]
\centering
\includegraphics[width=7cm]{sc_noisy.png}
\caption{Same image as Figure~\ref{fig:threech}, but with 'salt \& pepper' noise with a variation of 0.1. Decreasing $\lambda$ from 255*255 to 125 allowed less
noise to be detected. Size = 200 x 200, $\phi_{0}(x,y) = - \sqrt{(x - 100)^2 + (y - 100)^2)} +
50$, no reinitialization. Cpu = 3.17 sec and iterations = 5 when $\lambda = 255*255$. Cpu = 146.46 sec and iterations = 125 when $\lambda =
125$.}
\label{fig:sc_noisy}
\end{figure}

@@

Unfortunately, the complement did not work properly in our implementation. We closely followed the definitions in the original paper where $z_i^{in'} = 1 -
z_i^{in}$ and $z_i^{out'} = 1 - z_i^{out}$. The results was actually successfully detected, but continuous flickering occurred, and the curve did not stop in
one position. We discuss this results in more details in Section~\ref{sec:difficulties}.
